---
author: "Carl Boettiger"
output: 
  md_document:
    variant: markdown_github
---

```{r, message=FALSE}
library("dplyr")
library("tidyr")
library("ggplot2")
library("multipleuncertainty")
```

```{r}
devtools::install_github("rstudio/profvis")

profvis::profvis({
  devtools::load_all()

   grid <- seq(0, 200, by = 0.5)
  growth <- multiple_uncertainty(x_grid = grid, sigma_g = 0.5, sigma_m = 0.1, sigma_i = 0.1)
})
```


With parallel BLAS libraries enabled, parts of the multiple_uncertainty calculation are already parallelized.  
On machines with small numbers of cores, it might make sense to run the `multiple_uncertainty`

```{r}
library("parallel")

fig3 <- function(cost, dr, noise){  
  grid <- seq(0, 200, by = 0.5)
  price <- 1
  profit <- function(x,h) price * pmin(x,h) - cost * h - dr * h ^ 2
  
  
  o <- mclapply(
    list(small = c(g = 0.1, m = 0.1, i = 0.1),
         growth = c(g = 0.5, m = 0.1, i = 0.1),
         measure = c(g = 0.1, m = 0.5, i = 0.1),
         implement = c(g = 0.1, m = 0.1, i = 0.5)), 
    function(s)
      multiple_uncertainty(f = logistic, x_grid = grid, sigma_g = s$g, sigma_m = s$m, sigma_i = s$i, noise_dist = noise, profit_fn = profit))
  
  df <- data.frame(y_grid = grid, small = small, growth = growth, 
                   measure = measure, implement = implement) %>%
    tidyr::gather(scenario, value, -y_grid)
}
```




One option is to parallelize the `do` loop with `multidplyr`, however, this requires essentially `ncores` times as much memory, which is very inefficient.  

```{r eval=FALSE}
cl <- get_default_cluster()
cl %>% cluster_assign_value("fig3", fig3) %>% cluster_library("multipleuncertainty")

expand.grid(cost = c(0, 0.02, 0.2),
            dr = c(0, 0.1),
            noise = c("uniform", "lognormal")) %>%
  multidplyr::partition() %>%
  dplyr::group_by(cost, dr, noise) %>%
  dplyr::do(fig3(.$cost, .$dr, .$noise)) -> df

```



```{r}
df %>% 
  dplyr::filter(noise == "uniform") %>%
  ggplot(aes(x = y_grid, y = value, col = scenario)) + 
    geom_line()  + 
    facet_grid(cost ~ dr) + 
    xlab("Stock") + 
    ylab("Escapement") + 
    coord_cartesian(xlim = c(0, 150), ylim = c(0,100)) + 
    theme_bw()
```




## Reed Result


```{r}
reed <- multiple_uncertainty(noise_dist = "lognormal")

qplot(seq_along(reed), reed) + 
    coord_cartesian(xlim = c(0, 150), ylim = c(0, 150)) + 
    xlab("Stock") + ylab("Escapement") +
    theme_bw()
```
