---
title: "Fig3 robustness appendix"
author: "Carl Boettiger"
date: "9/24/2015"
output: 
  md_document:
    variant: markdown_github
---

```{r, message=FALSE}
library("dplyr")
library("tidyr")
library("ggplot2")
library("multipleuncertainty")
knitr::opts_chunk$set(cache = TRUE)
```

## Results are robust to grid

We confirm the general pattern observed is independent of the choice of grid step-size or the grid's maximum range.  To do so, we consider all combinations of maximum grid range and grid step size, using possible maximum ranges of `150`, `200`, `300`, and `400`, and possible step sizes of `0.5`, `1`, and `2`. Note that the largest grid thus has 800 grid points and the resulting linear algebra may need 16 GB of memory.

```{r}
fig3 <- function(max, by, noise="uniform"){  
  grid <- seq(0, max, by = by)
  small     <- multiple_uncertainty(f = logistic, x_grid = grid, sigma_g = 0.1, sigma_m = 0.1, sigma_i = 0.1, noise_dist = noise)
  growth    <- multiple_uncertainty(f = logistic, x_grid = grid, sigma_g = 0.5, sigma_m = 0.1, sigma_i = 0.1, noise_dist = noise)
  measure   <- multiple_uncertainty(f = logistic, x_grid = grid, sigma_g = 0.1, sigma_m = 0.5, sigma_i = 0.1, noise_dist = noise)
  implement <- multiple_uncertainty(f = logistic, x_grid = grid, sigma_g = 0.1, sigma_m = 0.1, sigma_i = 0.5, noise_dist = noise)
  df <- data.frame(y_grid = grid, small = small, growth = growth, 
                   measure = measure, implement = implement) %>%
    tidyr::gather(scenario, value, -y_grid)
}

df <- 
expand.grid(max = c(150, 200, 300, 400), 
            by = c(0.5, 1, 2)) %>%
  dplyr::group_by(max,by) %>%
  dplyr::do(fig3(.$max, .$by, "uniform"))

```



```{r fig.width = 9, fig.height = 12}
plt <- function(df)
  df %>% ggplot(aes(x = y_grid, y = value, col = scenario)) + 
    geom_line()  + 
    facet_grid(by ~ max) + 
    xlab("Stock") + 
    ylab("Escapement") + 
    coord_cartesian(xlim = c(0, 150), ylim = c(0,100)) + 
    theme_bw()

df %>% plt()
```

We can repeat this analysis for lognormal noise, which we find to be even less sensitive to the small jitter created at the coarser grid sizes.  


```{r}
df <- 
expand.grid(max = c(150, 200, 300, 400), 
            by = c(0.5, 1, 2)) %>%
  dplyr::group_by(max,by) %>%
  dplyr::do(fig3(.$max, .$by, "lognormal"))
```



```{r fig.width = 9, fig.height = 12}
df %>% plt()
```


Based on these observations, we selected a maximum of `200` and and a `delta` of `0.5` as our chosen grid for most of the analysis.


## Robustness to noise level

We consider the results across a range of possible values for the "low" and "high" noise limits chosen. 

```{r warning=FALSE}
fig3 <- function(low, high, noise_dist){  
  grid <- seq(0, 200, length = 401)
  model <- "logistic"
  small     <- multiple_uncertainty(f = model, x_grid = grid, sigma_g = low, sigma_m = low, sigma_i = low, noise_dist = noise_dist)
  growth    <- multiple_uncertainty(f = model, x_grid = grid, sigma_g = high, sigma_m = low, sigma_i = low, noise_dist = noise_dist)
  measure   <- multiple_uncertainty(f = model, x_grid = grid, sigma_g = low, sigma_m = high, sigma_i = low, noise_dist = noise_dist)
  implement <- multiple_uncertainty(f = model, x_grid = grid, sigma_g = low, sigma_m = low, sigma_i = high, noise_dist = noise_dist)
  ## Combine records by scenario
  df <- data.frame(y_grid = grid, small = small, growth = growth, 
                   measure = measure, implement = implement) %>%
    tidyr::gather(scenario, value, -y_grid)
}
```

```{r warning=FALSE}
expand.grid(low = c(0, 0.01, 0.1), high = c(0.2, 0.5, 0.8), noise_dist = c("lognormal", "uniform")) %>%
  dplyr::group_by(low, high, noise_dist) %>%
  dplyr::do(fig3(.$low, .$high, .$noise_dist)) -> df

```



```{r fig.width = 6, fig.height = 8}
df %>% filter(noise_dist == "lognormal") %>%
  ggplot(aes(x = y_grid, y = value, col = scenario)) + 
    geom_line()  + 
    facet_grid(high ~ low) + 
    xlab("Stock") + 
    ylab("Escapement") + 
    coord_cartesian(xlim = c(0, 150), ylim = c(0,100)) + 
    theme_bw() + 
    ggtitle("Lognormal noise")

```


```{r fig.width = 6, fig.height = 8}
df %>% filter(noise_dist == "uniform") %>%
  ggplot(aes(x = y_grid, y = value, col = scenario)) + 
    geom_line()  + 
    facet_grid(high ~ low) + 
    xlab("Stock") + 
    ylab("Escapement") + 
    coord_cartesian(xlim = c(0, 150), ylim = c(0,100)) + 
    theme_bw() + 
    ggtitle("Uniform noise")

```

## Robustness to model

We consider a variety of functional forms for the population growth model

Simulation of the growth function illustrates roughly comparable parameter values (e.g. same unfished equilibrium value)

```{r}
sims <- function(model, r){
  f <- get(as.character(model))
  x <- numeric(10)
  x[1] <- 60
  for(t in 1:length(x))
    x[t+1] = f(x[t], 0, r = r)
  data.frame(x, t = 1:length(x))
}
```


```{r}
options(stringsAsFactors = FALSE)

expand.grid(model = c("logistic", "bevertonholt", "ricker", "gompertz"), 
            r = c(0.1, 0.5, 0.75, 1)) %>%
  dplyr::group_by(model, r) %>%
  dplyr::do(sims(.$model, .$r)) -> 
  df

ggplot(df, aes(x = t, y = x, color = model)) + geom_line() + facet_grid(model ~ r)

```

